---
title: "sem3_spesification.R"
author: "E"
date: "1/27/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Seminar 3 OLS, Spesification

### 1)

Model: $$\ y = E[y|x,z] + \epsilon $$

Choose preferred specification, and the object is to predict.

```{r, hide = F}

sem1ex1 <- haven::read_dta("data/econ4137sem1ex1.dta")

summary(sem1ex1)

# 1) linaer
sem1ex1 %>%  lm( y ~ x , data = .) %>% summary()

# log -log
sem1ex1 %>%  lm( log(y) ~ I( log(x)) + z , data = .) %>% summary()

# log-log
sem1ex1 %>%  lm( log(y) ~ I( log(x)) , data = .) %>% summary()

```

#### ii) Test functional form

With two alternative models that are non nested, the R² and AIC/BIC cant be used. An alternativ approch is to use the Box-Cox transformation (PE-test).

*1) Estimate both models.*

*2) predict* $\hat{y}$ and $\hat{log(y)}$\*

\*3) t-statistics of $\delta$ and model $y = x\beta + \delta_{tLLM}[(log(y) - log(\hat{y})] + \epsilon$

```{r}

model_l <- sem1ex1 %>% lm( y ~x, data = .)
model_log <- sem1ex1 %>% lm( I(log(y)) ~I(log(x)), data = .)

# fit the model
sem1ex1$yhat <- model_l$fitted.values
sem1ex1$yhat_log <- model_log$fitted.values

PE_test <- sem1ex1 %>% 
  mutate( yhat_test = (log(yhat) - yhat_log),
          ) %>% 
  lm( y ~ x + yhat_test, data = .)

summary(PE_test)

```

PE_test - cant rejcet log model. Appear that linear model is preferred.

```{r}
PE_test2 <- sem1ex1 %>% 
  mutate( log_test = (yhat - exp(yhat_log) ) )%>% 
  lm( y ~ x + log_test, data = .)

summary(PE_test2)
```

cant reject. Seems that also log-model are preferred.

## 2 3.2 Verbeek

Clothing.dta contains information of sales, size and other characteristics of 400 Dutch men\`s fashion stores.

Goal is to explain sales per squaremetre (sales).

```{r}
fs::ls_dir()
clothing <- haven::read_dta("data/clothing.dta")

head(clothing)
```

#### a1)

The linear model D:

$$
log(sales) = x\beta + \epsilon  
$$

Where x = {log sales, log shop size and a constant}.

```{r}
mD <- clothing %>% mutate( lnsales = log(sales),
                           lnhourw = log(hoursw),
                           lnsize = log(ssize)
                           ) %>% 
  lm( lnsales ~ lnhourw + lnsize , data = .)

summary(mD)
```

#### a2) The linear model: $sales = \alpha + x\beta + hoursw + size$

```{r}

modelA <- clothing %>% lm( sales ~ hoursw + ssize, data = .)

summary(modelA)
```

#### b) RESET

```{r}
clothing$ma_fitted <- (modelA$fitted.values)^2

RESET_ma <- clothing %>% lm( sales ~ hourspw + ssize + ma_fitted, data = .)

summary(RESET_ma)

```

The RESET-test fail to reject \delta.

c) Does number of owner affects shop sales?

```{r}
model_owner <- clothing %>% lm( sales ~ hoursw + ssize + nown, data = .)

summary(model_owner)

# F-test
f_statistic <- function(mR, mUnR){
  
  R2rest <- summary(mR)$r.squared 
  R2unr <- summary(mUnR)$r.squared
    
  ((R2unr-R2rest)/1)/( (1-R2unr)/(nrow(summary(mR$model))-5))  
  
}

# Low F-stat -> cant reject H0, owner has impact
f_statistic(mR = modelA, mUnR = model_owner)

model_part_time <- clothing %>% lm( sales ~ hoursw + ssize + npart, data = .)

# Low F-stat -> cant reject H0, owner has impact
f_statistic(mR = modelA, mUnR = model_part_time)

```

#### e) Model B: sales ~ owner, full-times workers,parttime workers and shop size
```{r}

modelB <- clothing %>% lm( sales ~ nown + nfull + npart + ssize , data = .)


modelB %>% summary()


```

```{r}
summary(modelA)$adj.r.squared
summary(modelB)$adj.r.squared

# non-nested F-test
clothing$mA_fitted <- modelA$fitted.values
clothing$mB_fitted <- modelB$fitted.values

PE <- clothing %>%
  mutate( y_t = (sales - mB_fitted) ) %>% 
  lm( sales ~ hoursw + ssize + y_t , data = .)

PE2 <- clothing %>%
  mutate( y_t = (sales - mA_fitted) ) %>% 
  lm( sales ~ nown + nfull + npart + ssize + y_t , data = .)

summary(PE)
summary(PE2)
```


### h) The J-test
```{r}

```







