---
title: "Heteroskedasticity210130"
author: "(Lecturenotes and Verbeek chapter 4)"
date: "1/30/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Heteroskedasticity

-   Residual variance increases with the gap between regression line and CEF

Occurs when assumptions $f(x) = x\beta$ is wrong

```{r}
x  <- sample(c(1:100), size = 100, replace = T)
u <- rnorm(n = 100, mean = 0, sd = (x^1.5)^.5 )

df <- tibble( x = x, y = 1 + 0.5*x  +u)

model_wrong_f <- df %>% lm( y ~ x, data = .)


df$uhat <- model_wrong_f$residuals

```

Illustration of the hetero

```{r}
df %>% ggplot( aes( y = uhat, x = x)) + geom_point()
```

### Example 2) Wooldrigde price data

```{r}
hprice1 <- wooldridge::hprice1 %>% as_tibble()

library(estimatr)
model <- "price ~ lotsize + sqrft"

hprice1

# OLS
price_house_model <- hprice1 %>% lm( price ~ lotsize + sqrft , data = . ) 

#
summary(price_house_model)

# OLS with robust
hprice1 %>% estimatr::lm_robust( price ~ lotsize + sqrft, se_type = "HC1" , data = . ) %>% summary()

```

#### Testing for Heteroskedasticity

1) White Test: OLS, auxiliary regression equation

- i) \$\e² = \X\beta + \X\^2\\beta\_{2} + \\(X1\*X2B+..)\$

- ii) R² should be small. --\> Not only heterosked. thats tested byt also error

#### Example 1
```{r}
price_model_1 <- hprice1 %>% lm( price  ~ lotsize + sqrft , data = .)

hprice1$res <- resid(price_model_1)

hprice1 %>% lm( I(res^2) ~ lotsize + sqrft , data = .) %>% summary()

bptest(price_model_1)
bptest(price_model_1, ~ fitted(price_model_1) + I(fitted(price_model_1)) )

```



##### Example 2

```{r example_white_test}

price_model <- hprice1 %>%  lm( price ~ lotsize + sqrft + bdrms, data = . )

summary(price_house_model)



# BP-test
# Auxiliary : price ~ x + y + x² + y² + x*y
library(lmtest)
bptest( price_model)

broom::augment(price_model) %>% mutate( u2 = (.resid)^2) 
resid(price_model)^2 %>% head()

# Auxliary-eq
aux_reg <- broom::augment(price_model) %>% mutate( u2 = (.resid)^2) %>% 
  # Model
  lm( u2 ~ lotsize + sqrft + bdrms, data = .)

# White-test
bptest( price_model, ~ fitted(price_model) + I(fitted(price_model)^2) )
```

#### Example 3

```{r}
price_m3 <- as_tibble(hprice1) %>% lm( lprice ~ llotsize + sqrft, data = .) 

summary(price_m3)

# Robust se
as_tibble(hprice1) %>% estimatr::lm_robust( lprice ~ llotsize + sqrft, se_type = "HC1" ,data = .) 

## White
price_model_res_2 <- hprice1 %>%
  mutate( u2 = (resid(price_m3)^2) ) %>% 
  lm( u2 ~llotsize + sqrft + I(llotsize^2) + I(sqrft^2) +  I(sqrft*llotsize) , data  = .)
 
# 2
price_model_res <- hprice1 %>%
  mutate( u2 = (resid(price_m3)^2) ) %>% 
  lm( u2 ~llotsize + sqrft, data = .)

summary(price_model_res_2) 

(summary(price_model_res_2)$r.squared*nrow(hprice1))

library(lmtest)
bptest( price_m3, ~ fitted(price_m3) + I(fitted(price_m3)^2) )
bptest( price_m3, ~ llotsize + sqrft + I(llotsize^2) + I(sqrft^2) +  I(sqrft*llotsize), data = hprice1)
pchisq((summary(price_model_res)$r.squared*nrow(hprice1)), (2) ,lower.tail = FALSE)

## Alt 2

price_m4 <- hprice1 %>% lm( I(log(price)) ~ I(log(lotsize)) +I(log(sqrft)) + bdrms, data = . )

price_m4 %>% summary()

bptest(price_m4)
bptest(price_m4, ~ fitted(price_m4) + I(fitted(price_m4)^2) )

```







