---
title: "2021-02-04 Example heteroskedastecity"
author: "(by Verbeek (2012) )"
date: "2/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
theme_set(theme_light())

wage2 <- data.table::fread("data/data_verbeek/labour2.prn") %>% janitor::clean_names()



```

### Example Labour demand.
Data from Belgium, and 569 firms. The data is cross-sectional, from 1996. Average wage, amount of capital, and measure of output.
  -labour:total employment (number of workers);
  -capital :total fixed assets (in million euro);
  -wage:total wage costs divided by number of workers (in 1000 euro);
  -output:value added (in million euro).




Model: approximate r by the capital stock K

#### OLS
```{r}
model_1_L <- wage2 %>% lm( labour ~ capital + output + wage, data = .)

summary(model_1_L)
```


- 1) Test for hetro: Breusch-Pagan test:
```{r}

auxilar_model_1 <- tibble(e2 = residuals(model_1_L)^2, wage2) %>% 
  lm( e2 ~ capital + output + wage, data =.)

summary(auxilar_model_1)


```

Data contains some small and some large firms, values may diviate a lot, including unobserved characteristics in e.

##### 1.step: Use log-linear model.
```{r}
# Use log
wage2 %>% ggplot( aes( x = wage)) + geom_histogram() + scale_x_log10()

```

Model demand for L

```{r}
log_model_l_demand <- wage2 %>% 
  lm( I(log(labour)) ~I(log(wage)) + I(log(output)) + I(log(capital)),
      data = .)

summary(log_model_l_demand)
```

*-*Interpretation: 1% increaste output, leads to 0.99% increase in L-demand.

Test for heteroskedasticity:

```{r}
wage2$residuals_2 <- (log_model_l_demand$residuals)^2 


BP_test <- wage2 %>% 
  lm( residuals_2 ~I(log(wage)) + I(log(output)) + I(log(capital)),
      data = .) 

chi_test_score <-summary(BP_test)$r.squared*nrow(wage2)


pchisq(chi_test_score, df = 3, lower.tail = F)

```

Which is on margin above 5. J = regressors (equal = 3)

#### The White-test

*-* The Auxiliary regression white test:

```{r}

wage_2 <- wage2 %>% 
  mutate(
    log_wage = log(wage),
    log_output = log(output),
    log_capital = log(capital),
    log_w_2 = log_wage^2,
    log_out_2 = log_output^2,
    log_capital_2 = log_capital^2,
    l_w_l_o = log_wage*log_output,
    l_w_l_c = log_wage*log_capital,
    l_o_l_c = log_output*log_capital
  )  

wage_2 <-wage_2 %>%   lm( residuals_2 ~log_wage+log_output+log_capital+log_w_2+log_out_2+log_capital_2+
        l_w_l_o + l_w_l_c + l_o_l_c,      data = .) 





chi_sq_value_test <- summary(white_test)$r.squared*nrow(wage2) 

# Significant, can reject Hetroskedasteciut
pchisq(chi_sq_value_test, df = 9, lower.tail = F)

nrow(wage2)

```

*-* Report White-robust standard errors:

```{r}
l_model_robust_se <- wage2 %>% estimatr::lm_robust( I(log(labour)) ~I(log(wage)) + I(log(output)) + I(log(wage)) + I(log(capital)), se_type = "HC1" ,data = . )

summary(log_model_l_demand)
summary(l_model_robust_se)
```

### EGLS

```{r}
wage2$resudal_2 <- (log_model_l_demand$residual)^2
```

```{r}
e_model <-wage2 %>% 
  lm( I(log(resudal_2)) ~ I(log(wage)) + I(log(output)) + I(log(capital)),
      data = .)


summary(e_model)
```
The summary result lead to rejection of H0 of homosedasticy


```{r}

# Transfomr the fitted value exp(:)^0.5
wage2$exp_fitted_e <- (exp(e_model$fitted.values))^.5

model_egls <- wage2 %>% 
  mutate( log_l_e = log(labour)/exp_fitted_e,
          log_w_e = log(wage)/exp_fitted_e,
          log_o_e = log(output)/exp_fitted_e,
          log_c_e = log(capital)/exp_fitted_e,
          c_e = 1/exp_fitted_e
          ) %>% 
  lm( log_l_e ~ 0 + c_e + log_w_e + log_o_e  + log_c_e  , data = .)

summary(model_egls)
summary( log_model_l_demand)

cor(log(wage2$labour), (model_egls$fitted.values))^2
# Whcih is lower than OLS
```





