---
title: "2021-02-04 Example heteroskedastecity"
author: "(by Verbeek (2012) )"
date: "2/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
theme_set(theme_light())

wage2 <- data.table::fread("data/data_verbeek/labour2.prn") %>% janitor::clean_names()



```

### Example Labour demand.
Data from Belgium, and 569 firms. The data is cross-sectional, from 1996. Average wage, amount of capital, and measure of output.
  -labour:total employment (number of workers);
  -capital :total fixed assets (in million euro);
  -wage:total wage costs divided by number of workers (in 1000 euro);
  -output:value added (in million euro).




Model: approximate r by the capital stock K

#### OLS
```{r}
model_1_L <- wage2 %>% lm( labour ~ capital + output + wage, data = .)

summary(model_1_L)
```


- 1) Test for hetro: Breusch-Pagan test:
```{r}

auxilar_model_1 <- tibble(e2 = residuals(model_1_L)^2, wage2) %>% 
  lm( e2 ~ capital + output + wage, data =.)

summary(auxilar_model_1)


```

Data contains some small and some large firms, values may diviate a lot, including unobserved characteristics in e 
```{r}
# Use log
wage2 %>% ggplot( aes( x = wage)) + geom_histogram() + scale_x_log10()

```


